---
title: "Using Postgre with R"
output: html_notebook
---

First, we need to establish a connection with our database. For this, we will use the RPostgres and 

```{r}
library(DBI)
library(RPostgres)
library(tidyverse)
```

Now, we'll create our connection.
```{r}
con <- dbConnect(RPostgres::Postgres(),      # Use the postgres driver
                 dbname = 'prescribers',     # Name of the database we want to connect to
                 host = 'localhost',         # 'localhost' or eg. 'ec2-54-83-201-96.compute-1.amazonaws.com'
                 port = 5432, 
                 user = 'postgres',
                 password = 'postgres')
```

We can see the tables in the database.
```{r}
dbListTables(con)
```

We can even see the columns of a particular table.
```{r}
dbListFields(con, 'drug')
```

To execute a query, we can use the dbSendQuery function
```{r}
# proportion of opioid prescriptions (step 1)
query = "SELECT f.fipscounty, f.county , SUM(total_claim_count) as total, population
FROM drug AS d
INNER JOIN prescription AS p2
USING (drug_name)
INNER JOIN prescriber AS p1
USING (npi)
INNER JOIN zip_fips AS z
ON p1.nppes_provider_zip5 = z.zip
INNER JOIN fips_county as f
USING (fipscounty)
INNER JOIN population as p3
USING (fipscounty)
WHERE f.state = 'TN' AND opioid_drug_flag = 'Y'
GROUP BY f.fipscounty, f.county, population
ORDER BY total DESC;"

res <- dbSendQuery(con, query)
```

Then to retrieve the results, we can use dbFetch. Note that you can fetch only a portion of the results if needed, but we'll fetch everything.
```{r}
prescribers <- dbFetch(res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res)
```

The results that are returned come through as a dataframe, so we can manipulate it using tidyverse.
```{r}
# prescribers %>% 
#   filter(nppes_provider_city == 'NASHVILLE')

# oroportion (step 2)
prescribers1 <-prescribers %>%
  mutate(opioid_proportion = total / sum(total) * 100,
         pop_proportion = population / sum(population) * 100,
         opioid_per_pop = total / population) %>%
  arrange(desc(opioid_per_pop))

prescribers1 %>% 
  head(15) %>% 
  ggplot(aes(x=opioid_per_pop, y=reorder(county, opioid_per_pop))) + geom_col()

```
```{r}
# explorting overdose deaths by county
query = "SELECT SUM(overdose_deaths), fipscounty
FROM overdose_deaths
GROUP BY fipscounty;"

res <- dbSendQuery(con, query)
```

```{r}
overdose <- dbFetch(res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res)
```

```{r}
# prescribers %>% 
#   filter(nppes_provider_city == 'NASHVILLE')

#proportion (step 2)
overdose1 <-overdose %>%
  left_join(prescribers1, by = "fipscounty") %>% 
  mutate(prop_deaths = sum / population * 100)

overdose1%>% 
  ggplot(aes(x = opioid_per_pop, y = prop_deaths)) +
  geom_point() +
  scale_x_log10()
  
```
```{r}
cor(overdose1$opioid_per_pop, overdose1$prop_deaths)
```


```{r}
library(maps)
library(plotly)
library(colorspace)
library(tidyverse)
library(ggplot2)
# install.packages("ggiraph")
library(ggiraph)
```

```{r}
states <- map_data("state")

head(states)

states <- map_data("state")
tn<-subset(states, region %in% c("tennessee"))
tn_df<-subset(states, region=="tennessee")
counties<-map_data("county")
tn_county<-subset(counties, region=="tennessee")
tn_base <- ggplot(data = tn_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
tn_base +  
  geom_polygon(data = tn_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)

```
```{r}
# change column names to uppercase; then rename
tn_county[,6] = toupper(tn_county[,6])
tn_county <- tn_county %>% 
  rename(county = subregion) %>% 
  mutate(county = str_replace_all(county, " ", ""))

overdose1 <- overdose1 %>% 
  mutate(county = str_replace_all(county, " ", ""))

future_tibble <- tn_county %>% 
  inner_join(overdose1, by = "county")
  

```

```{r}
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())
```


```{r}
elb<-tn_base + 
  geom_polygon(data = future_tibble, aes(fill = prop_deaths, text = sprintf('County: %s\nDeath Proportion: %s', county,round(prop_deaths,2))), color = "black") +
  geom_polygon(color = "black") +
  theme_bw() +
  ditch_the_axes
```

```{r}
elb<-elb +
            scale_fill_gradientn(colours = colorspace::diverging_hcl(3,palette = "Blue-Red"))

```

```{r}
ggplotly(elb,tooltip = "text")
```

```{r}
# Question 3
q3_query = "SELECT *
FROM drug AS d
INNER JOIN prescription AS p2
USING (drug_name)
INNER JOIN prescriber AS p1
USING (npi)
INNER JOIN zip_fips AS z
ON p1.nppes_provider_zip5 = z.zip
INNER JOIN fips_county as f
USING (fipscounty)
INNER JOIN overdose_deaths as o
USING (fipscounty)
WHERE f.state = 'TN' AND opioid_drug_flag = 'Y' AND YEAR BETWEEN 2015 AND 2018
ORDER BY overdose_deaths DESC;"

q3_res <- dbSendQuery(con, q3_query)
```

```{r}
od_trend <- dbFetch(q3_res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(q3_res)
```

```{r}

```

```{r}

```

```{r}
# Question 5
query2 = "With CTE AS (
	SELECT
	fipscounty, zip, tot_ratio AS highest_ratio,
	RANK() OVER(
		PARTITION BY zip
		ORDER BY tot_ratio DESC)
	FROM zip_fips
	WHERE fipscounty LIKE '47%'
	ORDER BY zip
),
CTE2 AS (
	SELECT fipscounty, zip
	FROM CTE
	WHERE rank = 1)
SELECT p.npi, CTE2.zip, CTE2.fipscounty, fp.county, drug_name, generic_name, p.total_day_supply, p.total_claim_count, 
CASE WHEN d.generic_name LIKE '%%CODEINE%%' THEN 'CODEINE'
            WHEN d.generic_name LIKE '%%FENTANYL%%' THEN 'FENTANYL'
            WHEN d.generic_name LIKE '%%HYDROCODONE%%' THEN 'HYDROCODONE'
            WHEN d.generic_name LIKE '%%MORPHINE%%' THEN 'MORPHINE'
            WHEN d.generic_name LIKE '%%OXYCODONE%%' THEN 'OXYCODONE'
            WHEN d.generic_name LIKE '%%OXYMORPHONE%%' THEN 'OXYMORPHONE'
            WHEN d.generic_name LIKE '%%TRAMADOL%%' THEN 'TRAMADOL'
            ELSE 'other' END 
            AS opioid
FROM drug AS d
INNER JOIN prescription as p
USING (drug_name)
INNER JOIN prescriber as p2
USING (npi)
INNER JOIN CTE2
On p2.nppes_provider_zip5 = CTE2.zip
INNER JOIN fips_county AS fp
USING(fipscounty)
WHERE opioid_drug_flag = 'Y'
ORDER BY p.total_day_supply desc;"

res2 <- dbSendQuery(con, query2)
```

```{r}
drugs <- dbFetch(res2)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res2)
```

```{r}
query1 = "SELECT fipscounty, SUM(overdose_deaths) AS deaths_fipscounty, population
FROM overdose_deaths
INNER JOIN population
USING(fipscounty)
GROUP BY fipscounty, population;"

res1 <- dbSendQuery(con, query1)
```

```{r}
overdose_deaths <- dbFetch(res1)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res1)
```

```{r}
all_opioid_data <- overdose_deaths %>% 
  inner_join(drugs, by = "fipscounty") %>% 
  mutate(prop_deaths = deaths_fipscounty / population * 100)

```

```{r}


```

