---
title: "Using Postgre with R"
output: html_notebook
---

First, we need to establish a connection with our database. For this, we will use the RPostgres and 

```{r}
library(DBI)
library(RPostgres)
library(tidyverse)
options(scipen = 999)
```

Now, we'll create our connection.
```{r}
con <- dbConnect(RPostgres::Postgres(),      # Use the postgres driver
                 dbname = 'prescribers',     # Name of the database we want to connect to
                 host = 'localhost',         # 'localhost' or eg. 'ec2-54-83-201-96.compute-1.amazonaws.com'
                 port = 5432, 
                 user = 'postgres',
                 password = 'postgres')
```

We can see the tables in the database.
```{r}
dbListTables(con)
```

We can even see the columns of a particular table.
```{r}
dbListFields(con, 'drug')
```

To execute a query, we can use the dbSendQuery function
```{r}
query = "SELECT f.county , SUM(total_claim_count) as total, population
FROM drug AS d
INNER JOIN prescription AS p2
USING (drug_name)
INNER JOIN prescriber AS p1
USING (npi)
INNER JOIN zip_fips AS z
ON p1.nppes_provider_zip5 = z.zip
INNER JOIN fips_county as f
USING (fipscounty)
INNER JOIN population as p3
USING (fipscounty)
WHERE f.state = 'TN' AND opioid_drug_flag = 'Y'
GROUP BY f.county, population
ORDER BY total DESC;
"

res <- dbSendQuery(con, query)
```

Then to retrieve the results, we can use dbFetch. Note that you can fetch only a portion of the results if needed, but we'll fetch everything.
```{r}
prescribers <- dbFetch(res)

# It is good practice to clear the result after fetching what you need in order to free all resources associated with the result set. 
dbClearResult(res)
```

The results that are returned come through as a dataframe, so we can manipulate it using tidyverse.
```{r}
prescribers1 <-prescribers %>%
  mutate(opioid_proportion = total / sum(total) * 100,
         pop_proportion = population / sum(population) * 100,
         opioid_per_pop = total / population) %>%
  arrange(desc(opioid_per_pop))

prescribers1 %>% 
  head(15) %>% 
  ggplot(aes(x=opioid_per_pop, y=reorder(county, opioid_per_pop))) + geom_col()

```